---
title: "Initial"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: sandstone
---
<body style="background-color:MintCream;">

<h3 style="text-align:center;font-size:185%;">[Home](https://pbischoff3.github.io/)  |  [Conferences](https://pbischoff3.github.io/Conferences) | [R Projects](https://pbischoff3.github.io/Projects)   |   [Publications](https://pbischoff3.github.io/Publications)  |  [Medical Hours](https://pbischoff3.github.io/Hours)</h3>

<h4 style="text-align:center;font-size:170%;">[MD PhD](https://pbischoff3.github.io/Projects/MD_PhD) | [BIOL 3100 Final](https://pbischoff3.github.io/Projects/BIOL_3100) </h4>

<h5 style="text-align:center;font-size:150%;">[Initial](https://pbischoff3.github.io/Projects/MD_PhD/initial) | [Secondary](https://pbischoff3.github.io/Projects/secondary) </h5>


```{r LAbels, include=FALSE}
library(dplyr)
library(janitor)
library(stringr)
library(readr)
library(ggplot2)
library(leaflet)
library(sf)
library(viridis)
library(htmltools)
library(stringr)
library(reshape2)
library(plotly)
library(gapminder)


my_df <- read.csv("../../MD_PhD/cleaned.csv")
```


## {.tabset}

### Maps 
```{r echo=FALSE}
pal1 <- colorFactor(
  palette = 'Reds', domain = my_df$research_rank)
pal2 <- colorFactor(
  palette = 'Blues', domain = my_df$primary_care_rank)

leaflet(my_df) %>% addTiles() %>%
  addMarkers(lng = ~longitude,lat = ~latitude,popup = ~school,label = ~school) %>% 
  addProviderTiles(providers$CartoDB.Positron)
```

Here is a description of all the MD/PhD programs offered in the nation. There are 115 schools, and my initial impression is how densely populated the schools are back east. Not a ton of Western options. 
<br>



### Data {.tabset}

#### Total Admittance
Here is the top 30 Total MD Admittance according to school.
<br>
```{r echo=FALSE}
keeps <- c("short_name","admittance_year")
temp = my_df[keeps]

top15 <- temp %>% 
  arrange(desc(admittance_year)) %>% 
  slice(1:30)

top15

top15 %>% 
  ggplot(mapping = aes(x = short_name, y = admittance_year))+
  geom_point()+
  theme(axis.text.x = element_text(angle=60, hjust=1))
```
This is for medical students overall, not just MD/PhD students. It seems awfully toploaded. This is a good basis to understand where my odds of acceptance may be higher than in others.
<br>
Exploring admittance for MD/PhD students is important as well.


#### MD/PhD Admittance
Here is the top 30 MD/PhD admittance according to school.
<br>
```{r echo=FALSE}
keeps2 <- c("short_name","md_ph_d_year_last_3_years")
temp2 = my_df[keeps2]

top15 <- temp2 %>% 
  arrange(desc(md_ph_d_year_last_3_years)) %>% 
  slice(1:30)

top15

top15 %>% 
  ggplot(mapping = aes(x = short_name, y = md_ph_d_year_last_3_years))+
  geom_point()+
  theme(axis.text.x = element_text(angle=60, hjust=1))
```
<br>

Not many students admitted each year across the board. It may be interesting to see the percentage of MD/PhD students admitted versus MD students admitted. 

#### MD vs MD/PhD Admittance %

```{r echo=FALSE}
my_df$admitpercent <- ((my_df$md_ph_d_year_last_3_years/my_df$admittance_year)*100)

keeps2 <- c("short_name","admitpercent")
temp2 = my_df[keeps2]

top15 <- temp2 %>% 
  arrange(desc(admitpercent)) %>% 
  slice(1:30)

top15
```
<br>
This graph shows us the highest percentage of MD/PhD students per the total of those admitted to medical school. Again, sort of a highest odds, kind of thing. 

#### 40 Furthest Distance from Temple

```{r echo=FALSE}
keeps2 <- c("short_name","distance_to_temple_mi")
temp2 = my_df[keeps2]

top15 <- temp2 %>% 
  arrange(desc(distance_to_temple_mi)) %>% 
  slice(1:40)

top15

top15 %>% 
  ggplot(mapping = aes(x = short_name, y = distance_to_temple_mi))+
  geom_point()+
  theme(axis.text.x = element_text(angle=60, hjust=1))
```
<br>
The temple is entirely important to me, so let's see which medical schools are the furthest from the temple. 


#### 60 Closest Distance to Temple 
```{r echo=FALSE}
keeps2 <- c("short_name","distance_to_temple_mi")
temp2 = my_df[keeps2]

top15 <- temp2 %>% 
  arrange(distance_to_temple_mi) %>% 
  slice(1:60)

top15

top15 %>% 
  ggplot(mapping = aes(x = short_name, y = distance_to_temple_mi))+
  geom_point()+
  theme(axis.text.x = element_text(angle=75, hjust=1))
```
<br>
And this shows the ones that are closest to the temple. 

#### 40 Lowest Home Prices
```{r echo=FALSE}
keeps2 <- c("short_name","home")
temp2 = my_df[keeps2]

top15 <- temp2 %>% 
  arrange(home) %>% 
  slice(1:40)

top15

top15 %>% 
  ggplot(mapping = aes(x = short_name, y = home))+
  geom_point()+
  theme(axis.text.x = element_text(angle=60, hjust=1))
```
<br>
An interesting part of data, but not one that I think I need to use to make the decision of what schools to apply to. This would be important for a tertiary level of decision making, or even as a means to help decide after being accepted to multiple schools. 

#### 40 Highest Home Prices
```{r echo=FALSE}
keeps2 <- c("short_name","home")
temp2 = my_df[keeps2]

top15 <- temp2 %>% 
  arrange(desc(home)) %>% 
  slice(1:40)

top15

top15 %>% 
  ggplot(mapping = aes(x = short_name, y = home))+
  geom_point()+
  theme(axis.text.x = element_text(angle=60, hjust=1))
```
<br>
Again, not a decision maker, but something to be mindful of for the future. 

