---
title: "Unit Tests"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: sandstone
---
<body style="background-color:MintCream;">

<h3 style="text-align:center;font-size:185%;">[Home](https://pbischoff3.github.io/)  |  [Conferences](https://pbischoff3.github.io/Conferences) | [R Projects](https://pbischoff3.github.io/Projects)   |   [Publications](https://pbischoff3.github.io/Publications)  |  [Medical Hours](https://pbischoff3.github.io/Hours)</h3>

<h4 style="text-align:center;font-size:150%;">[MD PhD](https://pbischoff3.github.io/Projects/MD_PhD) | [BIOL 3100 Final](https://pbischoff3.github.io/Projects/BIOL_3100) | [SPUR Preparation](https://pbischoff3.github.io/Projects/SPUR_Prep) </h4>

<h4 style="text-align:center;font-size:130%;">[netplot Overview](https://pbischoff3.github.io/Projects/SPUR_Prep/netplot) | [R Package Creation](https://pbischoff3.github.io/Projects/SPUR_Prep/package) | [Unit Tests](https://pbischoff3.github.io/Projects/SPUR_Prep/unit) | [Docker](https://pbischoff3.github.io/Projects/SPUR_Prep/docker) </h4>

# Walkthrough 1- "TestThat Package"
For this walk through, I followed the information given at [this website](https://www.geeksforgeeks.org/unit-testing-in-r-programming/).
<br>

First, we need to load the "testthat" package:
```{r}
library(testthat)
```
<br>

Next, we created a recursive program that calculates the factorial of a number 'n'. 
```{r}
factorial <- function(n)
{
  if(n == 0)
  {
    return(1)
  }
  else
  {
    return(n * factorial(n-2))
  }
}
```
<br>

Alright, now some unit testing. Is our function accurate?
```{r}
# import testthat package
library(testthat)

# use expect_that to create tests

```
Because of the failure of running this code, we can assume that the
