---
title: "ASMCUE 2023"
---

# Medical Experiences’ influence on Gender and PEERs 

Welcome to the website for our poster! It is here we will go through some of the traditional aspects of a poster, including the "Introduction", "Methods", "Results", and "Conclusion". If you have any questions, please do not hesitate to reach out to Porter at [portergbischoff\@gmail.com](mailto:portergbischoff@gmail.com){.email} !

## Abstract

Prior research has investigated the recruitment and retention of certain student identities (like gender (Master 2021), person excluded because of ethnicity or race (PEER) (Museus et al., 2010), and first-generation students (Bettencourt et al., 2020)) in STEM. No research has investigated the impact of having medical experiences on these groups. As such, our research questions were: What percentage of students indicate that they have medical experiences (Research Question #1)? Are medical experiences a benefit or barrier to students connecting and engaging with science? Specifically, do students who have medical experiences have higher or lower science motivation, sense of belonging, and science relevance (Research Question #2)?  

Data were collected from 366 students who were taking biology classes ranging from introductory biology for non-majors to advanced biology. All data were collected in a pre-survey/post-survey format. Students were asked if they had a medical experience, defined as: “Have you ever had a medical experience (hospitalization, chronic medical condition, and/or issues with mental, physical, emotional health etc.)?” Surprisingly, 57% of students surveyed reported having a medical experience and 22% reported having a chronic condition, indicating these are indeed an important and large identity to investigate further.  At the end of the semester, all students had significantly higher science identity, interest in science, sense of belonging, and value of the relevance of science. However, having medical experiences did not relate to consistent changes in constructs for all students. In contrast, consistent increases in constructs were found for first generation students with medical experiences such as: career value, interest, actions to promote success, self-efficacy, and sense of belonging. Information on the impact medical experiences can have on students from different demographics is crucial to investigate as science instructors modify course content and pedagogy to be more inclusive of all identities in the student population. 

## Author Information

Porter Bischoff\*, Kody Garrett, Clayton Rawson, Joshua Premo, Brittney N. Wyatt

Department of Biology, Utah Valley University, Orem, UT, 84058, USA.

\*Corresponding author: [portergbischoff\@gmail.com](mailto:portergbischoff@gmail.com){.email}

Acknowledgements: URSCA grant, NSF grant

Conflicts of Interest: The authors have no conflicts of interest to declare.

# Background



## Motivation for the Project

Porter has had a slew of medical experiences, and these experiences led him to pursue a career path in medicine. While going through his undergraduate experience, preparing for medical school, he found a real passion for science as well as medicine. Because this was relatively late in his education, he wishes that he knew this passion he had before his undergraduate studies. Therefore, this project was born, hoping to find that Porter wasn't uncommon in his experience and assisting other students understand their passions for science sooner than later. To learn more about how this project became to be, [please go to this link!](https://youtu.be/ZyujRGODbs4)

# Methods

We used an adapted survey from other literature to find information regarding STEM motivation. This survey was given pre-semester and post-semester to 14 different Biology courses at Utah Valley University. These classes had students who were both STEM majors and non-majors, as these classes were generally General Education classes. Utah Valley University is an open enrollment institution, making this data an interesting set to explore.

390 students answered the pre/post-semester surveys. We have a large number of pre-semester data, but only 390 responded to both the beginning and end of the semester survey.

## Aspects of the Survey

### STEM Motivation Questions

Here are the questions that we asked to determine the different aspects of STEM motivation.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(kableExtra)

# Define the data
data <- data.frame(
  Category = c("Science Identity", "Science Identity", "Science Identity", "Science Identity", "Science Identity",
               "Career Motivation", "Career Motivation", "Career Motivation", "Career Motivation", "Career Motivation",
               "Instrinsic Motivation", "Instrinsic Motivation", "Instrinsic Motivation", "Instrinsic Motivation", "Instrinsic Motivation",
               "Self-Determination", "Self-Determination", "Self-Determination", "Self-Determination", "Self-Determination",
               "Self-Efficacy", "Self-Efficacy", "Self-Efficacy", "Self-Efficacy",
               "Grade Motivation", "Grade Motivation", "Grade Motivation", "Grade Motivation", "Grade Motivation",
               "Value of Peer Interactions", "Value of Peer Interactions", "Value of Peer Interactions", "Value of Peer Interactions", "Value of Peer Interactions",
               "Sense of Belonging", "Sense of Belonging", "Sense of Belonging", "Sense of Belonging", "Sense of Belonging", "Sense of Belonging",
               "Competency in Science", "Competency in Science", "Competency in Science", "Competency in Science",
               "Community in Science", "Community in Science", "Community in Science", "Community in Science"
  ),
  Statement = c("I see myself as a biology person.", "I see myself as a science person.", "I see myself as a scientist.",
                "Other people (friends/family) see me as a science person.", "My science instructors see me as a science person.",
                "Learning science will help me get a good job.", "Understanding science will benefit me in my career.",
                "Knowing science will give me a career advantage.", "I will use science problem-solving skills in my career.",
                "My career will involve science.", "Learning science is interesting.", "I am curious about discoveries in science.",
                "The science I learn is relevant to my life.", "Learning science makes my life more meaningful.", "I enjoy learning science.",
                "I study hard to learn science.", "I prepare well for science tests and labs.", "I put enough effort into learning science.",
                "I spend a lot of time learning science.", "I use strategies to learn science well.", "I am confident that I will do well on science labs and projects.",
                "I believe I can earn a grade of 'A' in science.", "I am confident I will do well on science tests.",
                "I believe I can master science knowledge and skills.", "Scoring high on science tests and labs matters to me.",
                "It is important that I get an 'A' in science.", "I think about the grade I will get in science.",
                "Getting a good science grade is important to me.", "I like to do better than other students on science tests.",
                "My college science classes are more enjoyable when I work with other students.",
                "I understand more when my classmates participate in college classroom discussions.",
                "I learn best in college science classes when I work with classmates.",
                "The amount I understand is increased by classmates' ideas in my college science classes.",
                "When classmates share their ideas in college science classes, this helps me learn.",
                "When I am in a science setting, I feel a connection with the science community.",
                "When I am in a science setting, I feel respected.",
                "When I am in a science setting, I feel comfortable.",
                "When I am in a science setting, I trust my instructors to be committed to helping me learn.",
                "When I am in a science setting, I enjoy being an active participant.",
                "When I am in a science setting, I try to say as little as possible.",
                "I am good at solving complicated scientific problems.",
                "I read websites articles or books about scientific issues.",
                "I work on a project involving scientific concepts.",
                "I think about how my life is affected by science.",
                "To what extent is serving the community important to scientific work?",
                "To what extent is working with people important to scientific work?",
                "To what extent is helping others important to scientific work?",
                "To what extent is connecting with others important to scientific work?"
  )
)

# Create the table
table <- kable(data, format = "html", row.names = FALSE) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

# Display the table
table
```


## Statistical Analysis

We used Pearson Correlations to properly understand the significance data for each of the aspects. We are excited to continue doing analysis on this data set, as we know that we can dive deeper into it.

## Technology

SPSS was used to complete the Pearson Correlation. R packages were used, with `corrplot` to create the correlation plot, `waffle` to create the demographic plots, and `ggplot2`, `dplyr`, and `gridExtra` to create the pre/post semester change plot.

# Results

To see the R code for this, [follow this link](https://github.com/pbischoff3/pbischoff3.github.io/blob/main/Projects/2023_saber/reference/)! <br>

<small> *Each block represents 1%, but there are some demographics that have less than 1%. These are identified by 1 block, though they are smaller than the 1%.* </small>

## Demographics

### Major

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load required library
library(waffle)

# Define the data
career_aspiration <- c(
  "Biology" = 109,
  "Non-biology STEM" = 59,
  "Pre-Professional" = 80,
  "Non-STEM" = 142
)

# Calculate the adjusted counts based on a 10x10 grid
counts <- round(career_aspiration / sum(career_aspiration) * 100)

# Define the color palette for each career aspiration
pal <- c(
  "Biology" = "#275d38",
  "Non-biology STEM" = "#8BD8F0",
  "Pre-Professional" = "#A6D19F",
  "Non-STEM" = "#074EAB"
)

# Create the waffle chart with distinct colors
chart <- waffle(counts,
                rows = 10,  # Number of rows in the chart
                colors = pal,  # Colors for each category
                size = 0.7) +  # Size of each waffle block
  theme(legend.position = "bottom")

chart
```

<br>

The data indicates the distribution of career aspirations among the respondents. Approximately 29% of them expressed an interest in Biology, while Non-biology STEM fields accounted for around 16% of the responses. Pre-Professional aspirations represented approximately 21% of the respondents, and Non-STEM career paths comprised about 34% of the total aspirations reported.

<br>

### Year in School

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load required library
library(waffle)

# Define the data
class_year <- c(
  "Freshman" = 132,
  "Sophomore" = 112,
  "Junior" = 66,
  "Senior" = 80
)

# Calculate the adjusted percentages to have a total of 100 squares
total_squares <- 100
adjusted_percentages <- class_year / sum(class_year) * total_squares

# Round down the adjusted percentages to integers
counts <- floor(adjusted_percentages)

# Calculate the difference between the desired total squares and the current count
remaining_squares <- total_squares - sum(counts)

# Distribute the remaining squares to the categories with the highest fractional part
frac_parts <- adjusted_percentages - counts
sorted_indices <- order(frac_parts, decreasing = TRUE)
counts[sorted_indices[1:remaining_squares]] <- counts[sorted_indices[1:remaining_squares]] + 1

# Define the color palette for each class year
pal <- c(
  "Freshman" = "#275d38",
  "Sophomore" = "#8BD8F0",
  "Junior" = "#A6D19F",
  "Senior" = "#074EAB"
)

# Create the waffle chart with distinct colors
chart <- waffle(counts,
                rows = 10,  # Number of rows in the chart
                colors = pal,  # Colors for each category
                size = 0.7) +  # Size of each waffle block
              theme(legend.position = "bottom")  # Move legend to the bottom of the graph

chart
```

<br>

The data represents the distribution of students across different academic years. Among the students surveyed, 36% were Freshmen, 31% were Sophomores, 18% were Juniors, and the remaining 22% were Seniors.

<br>

### Career Aspiration

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(waffle)

# Define the data
career_aspiration <- c(
  "Doctor" = 51,
  "Dentist" = 28,
  "Health Care Professional" = 80,
  "STEM Professional" = 77,
  "Non-STEM" = 154
)

# Calculate the adjusted counts based on a 10x10 grid
counts <- round(career_aspiration / sum(career_aspiration) * 100)

# Define colors for each career aspiration
career_colors <- c(
  "Doctor" = "#275d38",
  "Dentist" = "#8BD8F0",
  "Health Care Professional" = "#A6D19F",
  "STEM Professional" = "#A391CF",
  "Non-STEM" = "#074EAB"
)

# Create the waffle chart with distinct shades of green
chart <- waffle(counts,
                rows = 10,  # Number of rows in the chart
                colors = career_colors,  # Colors for each category
                size = 0.7) +  # Size of each waffle block
  theme(legend.position = "bottom")  # Move legend to the bottom of the graph

chart
```

<br>

Non-STEM took the largest percent of the graph with 39%. STEM professionals took nearly twenty percent (20%), which was similar to the 21% of those who wished to be health care professionals. Those who wish to be a dentist or a doctor took a combined 20% (with those who wish to be dentists taking 7% and those who wished to be doctors taking 13%).

<br>

### Ethnicity

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load required library
library(waffle)

# Define the data
ethnicity_data <- c(
  "Black" = 8,
  "Native American" = 3,
  "Middle Eastern" = 1,
  "Asian" = 19,
  "Latina/o" = 32,
  "Multiracial" = 8,
  "Pacific Islander" = 5,
  "White" = 314
)

# Calculate the adjusted percentages to have a total of 100 squares
total_squares <- 100
adjusted_percentages <- ethnicity_data / sum(ethnicity_data) * total_squares

# Round down the adjusted percentages to integers for categories with more than 1%
counts <- floor(adjusted_percentages[adjusted_percentages >= 1])

# Calculate the remaining squares for categories with less than 1%
remaining_squares <- total_squares - sum(counts)

# Identify categories with less than 1% representation and allocate one square to each
counts_below_1_percent <- adjusted_percentages[adjusted_percentages < 1]
counts_below_1_percent <- ifelse(counts_below_1_percent > 0, 1, 0)
remaining_squares <- remaining_squares - sum(counts_below_1_percent)

# Distribute the remaining squares to categories with more than 1% representation
counts <- c(counts, counts_below_1_percent)
frac_parts <- adjusted_percentages[adjusted_percentages >= 1] - counts[adjusted_percentages >= 1]
sorted_indices <- order(frac_parts, decreasing = TRUE)
counts[sorted_indices[1:remaining_squares]] <- counts[sorted_indices[1:remaining_squares]] + 1

# Define the color palette for each ethnicity category
pal <- c(
  "Black" = "#275d38",
  "Native American" = "#8BD8F0",
  "Middle Eastern" = "#A6D19F",
  "Asian" = "#A391CF",
  "Latina/o" = "#074EAB",
  "Multiracial" = "#BA56CC",
  "Pacific Islander" = "#25B81D",
  "White" = "#76A398"
)

# Create the waffle chart with distinct colors
chart <- waffle(counts,
                rows = 10,  # Number of rows in the chart
                colors = pal,  # Colors for each category
                size = 0.7) +  # Size of each waffle block
  theme(legend.position = "bottom")  # Move legend to the bottom of the graph

chart
```

<br>

The data illustrates the racial and ethnic distribution of the surveyed individuals. Among the respondents, 1.6% identified as Black, 0.6% as Native American, 0.2% as Middle Eastern, 3.8% as Asian, 6.4% as Latina/o, 1.6% as Multiracial, 1% as Pacific Islander, and the majority, 78.7%, identified as White.

<br>

### Gender

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load required library
library(waffle)

# Define the data
gender_data <- c(
  "Men" = 201,
  "Transgender" = 2,
  "Women" = 182,
  "Fluid / Gender Non-Conforming" = 5
)

# Calculate the adjusted percentages to have a total of 100 squares
total_squares <- 100
adjusted_percentages <- gender_data / sum(gender_data) * total_squares

# Round down the adjusted percentages to integers for categories with more than 1%
counts <- floor(adjusted_percentages[adjusted_percentages >= 1])

# Calculate the remaining squares for categories with less than 1%
remaining_squares <- total_squares - sum(counts)

# Identify categories with less than 1% representation and allocate one square to each
counts_below_1_percent <- adjusted_percentages[adjusted_percentages < 1]
counts_below_1_percent <- ifelse(counts_below_1_percent > 0, 1, 0)
remaining_squares <- remaining_squares - sum(counts_below_1_percent)

# Distribute the remaining squares to categories with more than 1% representation
counts <- c(counts, counts_below_1_percent)
frac_parts <- adjusted_percentages[adjusted_percentages >= 1] - counts[adjusted_percentages >= 1]
sorted_indices <- order(frac_parts, decreasing = TRUE)
counts[sorted_indices[1:remaining_squares]] <- counts[sorted_indices[1:remaining_squares]] + 1

# Define the color palette for each gender category
pal <- c(
  "Men" = "#275d38",
  "Transgender" = "#8BD8F0",
  "Women" = "#A6D19F",
  "Fluid / Gender Non-Conforming" = "#074EAB"
)

# Create the waffle chart with distinct colors
chart <- waffle(counts,
                rows = 10,  # Number of rows in the chart
                colors = pal,  # Colors for each category
                size = 0.7) +  # Size of each waffle block
  theme(legend.position = "bottom")  # Move legend to the bottom of the graph

chart
```

<br>

The data represents the gender distribution of the respondents. Among those surveyed, 50.8% identified as Men, 0.5% as Transgender, 46% as Women, and 1.3% as Fluid / Gender Non-Conforming.

<br>

### Religious?

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load required library
library(waffle)

# Define the data
religion_data <- c(
  "Non-Religious" = 106,
  "Religious" = 284
)

# Calculate the adjusted percentages to have a total of 100 squares
total_squares <- 100
adjusted_percentages <- religion_data / sum(religion_data) * total_squares

# Round down the adjusted percentages to integers for categories with more than 1%
counts <- floor(adjusted_percentages[adjusted_percentages >= 1])

# Calculate the remaining squares for categories with less than 1%
remaining_squares <- total_squares - sum(counts)

# Identify categories with less than 1% representation and allocate one square to each
counts_below_1_percent <- adjusted_percentages[adjusted_percentages < 1]
counts_below_1_percent <- ifelse(counts_below_1_percent > 0, 1, 0)
remaining_squares <- remaining_squares - sum(counts_below_1_percent)

# Distribute the remaining squares to categories with more than 1% representation
counts <- c(counts, counts_below_1_percent)
frac_parts <- adjusted_percentages[adjusted_percentages >= 1] - counts[adjusted_percentages >= 1]
sorted_indices <- order(frac_parts, decreasing = TRUE)
counts[sorted_indices[1:remaining_squares]] <- counts[sorted_indices[1:remaining_squares]] + 1

# Define the color palette for each religion category
pal <- c(
  "Non-Religious" = "#275d38",
  "Religious" = "#8BD8F0"
)

# Create the waffle chart with distinct colors
chart <- waffle(counts,
                rows = 10,  # Number of rows in the chart
                colors = pal,  # Colors for each category
                size = 0.7) +  # Size of each waffle block
  theme(legend.position = "bottom")  # Move legend to the bottom of the graph

chart
```

<br>

The data indicates the distribution of respondents based on their religious affiliations. Among those surveyed, approximately 27% identified as Non-Religious, while around 73% identified as Religious.

<br>

### First Generation Student?

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load required library
library(waffle)

# Define the data
generation_data <- c(
  "Continuing Generation Student" = 290,
  "First Generation Student" = 93
)

# Calculate the adjusted percentages to have a total of 100 squares
total_squares <- 100
adjusted_percentages <- generation_data / sum(generation_data) * total_squares

# Round down the adjusted percentages to integers for categories with more than 1%
counts <- floor(adjusted_percentages[adjusted_percentages >= 1])

# Calculate the remaining squares for categories with less than 1%
remaining_squares <- total_squares - sum(counts)

# Identify categories with less than 1% representation and allocate one square to each
counts_below_1_percent <- adjusted_percentages[adjusted_percentages < 1]
counts_below_1_percent <- ifelse(counts_below_1_percent > 0, 1, 0)
remaining_squares <- remaining_squares - sum(counts_below_1_percent)

# Distribute the remaining squares to categories with more than 1% representation
counts <- c(counts, counts_below_1_percent)
frac_parts <- adjusted_percentages[adjusted_percentages >= 1] - counts[adjusted_percentages >= 1]
sorted_indices <- order(frac_parts, decreasing = TRUE)
counts[sorted_indices[1:remaining_squares]] <- counts[sorted_indices[1:remaining_squares]] + 1

# Define the color palette for each generation category
pal <- c(
  "Continuing Generation Student" = "#275d38",
  "First Generation Student" = "#8BD8F0"
)

# Create the waffle chart with distinct colors
chart <- waffle(counts,
                rows = 10,  # Number of rows in the chart
                colors = pal,  # Colors for each category
                size = 0.7) +  # Size of each waffle block
  theme(legend.position = "bottom")  # Move legend to the bottom of the graph

chart
```

<br>

The data represents the distribution of students based on their generational status. Among those surveyed, approximately 76% identified as Continuing Generation Students (at least one parent attended college), while around 24% identified as First Generation Students (no parents attended college).

<br>

### Medical Experience?

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(waffle)

# Define the data
medical_experience <- c(
  "No" = 167,
  "Yes" = 223
)

# Calculate the counts based on a 10x10 grid
counts <- round(medical_experience / sum(medical_experience) * 100)

# Define colors for each category
experience_colors <- c(
  "No" = "#8BD8F0",
  "Yes" = "#275d38"
)

# Create the waffle chart
chart <- waffle(counts,
                rows = 10,  # Number of rows in the chart
                colors = experience_colors,  # Colors for each category
                size = 0.7) +  # Size of each waffle block
  theme(legend.position = "bottom")  # Move legend to the bottom of the graph

chart
```

<br>

The data illustrates the percentage of those who identified to have had a medical experience. Among those surveyed, approximately 44% answered "No" while around 56% answered "Yes".

<br>

### Chronic Condition?

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load required library
library(waffle)

# Define the data
response_data <- c(
  "No" = 127,
  "Yes" = 86
)

# Calculate the adjusted percentages to have a total of 100 squares
total_squares <- 100
adjusted_percentages <- response_data / sum(response_data) * total_squares

# Round down the adjusted percentages to integers for categories with more than 1%
counts <- floor(adjusted_percentages[adjusted_percentages >= 1])

# Calculate the remaining squares for categories with less than 1%
remaining_squares <- total_squares - sum(counts)

# Identify categories with less than 1% representation and allocate one square to each
counts_below_1_percent <- adjusted_percentages[adjusted_percentages < 1]
counts_below_1_percent <- ifelse(counts_below_1_percent > 0, 1, 0)
remaining_squares <- remaining_squares - sum(counts_below_1_percent)

# Distribute the remaining squares to categories with more than 1% representation
counts <- c(counts, counts_below_1_percent)
frac_parts <- adjusted_percentages[adjusted_percentages >= 1] - counts[adjusted_percentages >= 1]
sorted_indices <- order(frac_parts, decreasing = TRUE)
counts[sorted_indices[1:remaining_squares]] <- counts[sorted_indices[1:remaining_squares]] + 1

# Define the color palette for each response category
pal <- c(
  "No" = "#8BD8F0",
  "Yes" = "#275d38"
)

# Create the waffle chart with distinct colors
chart <- waffle(counts,
                rows = 10,  # Number of rows in the chart
                colors = pal,  # Colors for each category
                size = 0.7) +  # Size of each waffle block
  theme(legend.position = "bottom")  # Move legend to the bottom of the graph

chart
```

<br>

Of those who identified as had a medical experience, approximately 60% answered "No" while around 40% answered "Yes" to whether they have a chronic condition.

## PEER Status & Medical Conditions


## Gender & Medical Conditions



# Conclusion


## Negative Effects


## Positive Effects


## Importance


## Next Steps

